Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-15 paper/main.tex
QA to 5.3.3

\subsection{Chosen implementations}
\label{sec:chosen_implementations}
\input{tables/sha3_implementations}

@@

Although we believe the \ac{SUPERCOP} toolkit and results give good indication
of how fast the different candidate implementations are, we do have some
general comments on the way it works and how it relates to our results.

\paragraph{Implementations does not conform to \ac{NIST} specification} for
many of the candidates. By this we mean that there is no {\verb Update() }
function available for all implementations, and that \ac{SUPERCOP} only
measures speed of hashing full byte strings. We are questioning if this
discriminates implementations which actually can do more of their processing on
a call to {\verb Update() }. This is also a hassle for a programmer who wants
to use the implementations found in \ac{SUPERCOP}

\paragraph{Incorrect implementations} are present in \ac{SUPERCOP}. By this we
mean implementations that does not conform to the \acp{KAT} supplied or
implementations which fail completely as hash functions, see section
\ref{sec:choosen_implementations} for details on the JH implementation.

\paragraph{Extra compiler flags}
We do not compile the candidate implementation with the exact same arguments
that is used in \ac{SUPERCOP}, most notably we append the -fPIC flag. This flag
creates \emph{position independent code}, in other words machine instruction
code that can execute properly regardless of where in memory it resides. This
will make one less register available on the x86 architecture. While we cannot
directly prove that this will affect performance of the implementations. We do
however suspect it. A person writing an optimization for \ac{SUPERCOP} will have
to balance his use of resources, among them registers, to achieve maximum speed.
The loss of one register will then give a performance penalty.

@@

\subsection{Small files}
The results for small files, 1KB and 1KB, are presented in Figure
\ref{fig:graph:1001b} and \ref{fig:graph:1001kb}. We can clearly see that in
this case the \ac{SHA}-256 implementation used by Tahoe-\ac{LAFS} outperforms
our implementations, which is somewhat unexpected given the \ac{SUPERCOP}
results. However, if we take into account the overhead created by our bindings
to the underlying implementations (Table \ref{tbl:cython:overhead}), some of the
\ac{SHA}-3 candidate implementations actually outperform double \ac{SHA}-256,
which is expected. We do not know how much overhead the bindings for
\ac{SHA}-256 are causing, but we can safely assume that the overhead is less
than for our bindings. If we can eliminate all time spent in our bindings both
BLAKE and \ac{BMW} outperform double \ac{SHA}-256 in every measurement, with
Luffa, Shabal and Simd outperforming double \ac{SHA}-256 in most of them.

We also measured single \ac{SHA}-256 for these files, which we believe should
show a difference. The lengths of processed by the hash function are dominated
by sized of 106B for the 1B files, and also by 63B for the 1KB files (Table
\ref{tbl:tahoe:datauploaded} and \ref{tbl:tahoe:datadownloaded}. Even by the
best-case performance scenario from Table \ref{tbl:supercop:sha256d} we would
expect to see a performance penalty. However this performance penalty is so
small that it is typically hidden due to system noise.

\subsection{Large files}
For large files, 1MB, 100MB and 1GB, we can see that our implementations are
outperforming double \ac{SHA}-256. Also for the large files we can see there is
a performance penalty by our bindings (Table \ref{tbl:cython:overhead}),
however the relative overhead cause are small compared to the time spent in the
underlying implementations. The reason for this is that we are hashing larger
sizes of data, especially for the 100MB and 1GB files (Tables
\ref{tbl:tahoe:datauploaded} and \ref{tbl:tahoe:datauploaded}). For these
larger sizes the overhead caused by using double \ac{SHA}-256 opposed to single
is also relatively small and does not cause a noteworthy performance penalty.

>>>>

\subsection{Chosen implementations}
\label{sec:chosen_implementations}

\input{tables/sha3_implementations}

@@

Although we believe the \ac{SUPERCOP} toolkit results give good indication of
how fast the different implementations of the candidates are, we do have some
general comments on the way it works and how it relates to our results.

\paragraph{Implementations do not conform to \ac{NIST} specification} for
many of the candidates. By this we mean that there is no {\verb update() }
function available for all implementations, and that \ac{SUPERCOP} only
measures speed of hashing full byte strings. We are questioning if this
discriminates implementations which actually can do more of their processing on
a call to {\verb update() }. This issue is also a hassle for a programmer who
wants to use the implementations found in \ac{SUPERCOP}.

\paragraph{Incorrect implementations} are present in \ac{SUPERCOP}. By this we
mean implementations that does not conform to the \acp{KAT} supplied or
implementations which fail completely as hash functions, see
\autoref{sec:chosen_implementations} for details on the JH implementation.

\paragraph{Extra compiler flags.}

We do not compile the candidate implementation with the exact same arguments
that is used in \ac{SUPERCOP}, most notably we append the {\verb -fPIC } flag.
This flag creates \emph{position independent code}, in other words machine
instruction code that can execute properly regardless of where in memory it
resides. This will make one less register available on the x86 architecture.
While we cannot directly prove that this will affect performance of the
implementations, we do however suspect it. A person writing an optimization for
\ac{SUPERCOP} will have to balance his use of resources, among them registers,
to achieve maximum speed.  The loss of one register will therefore give a
performance penalty.

@@

\subsection{Small files}

The results for small files, 1B and 1KB, are presented in Figures
\ref{fig:graph:1001b} and \ref{fig:graph:1001kb}. We can clearly see that in
this case, the \ac{SHA}-256 implementation used by Tahoe-\ac{LAFS} outperforms
our implementations, which is somewhat unexpected given the \ac{SUPERCOP}
results. However, if we take into account the overhead created by our bindings
to the underlying implementations (Table \ref{tbl:cython:overhead}), some of the
\ac{SHA}-3 candidate implementations actually outperform double \ac{SHA}-256,
which is expected. We do not know how much overhead the bindings for
\ac{SHA}-256 are causing, but we can safely assume that the overhead is far less
than for our bindings. If we can eliminate all extra time spent in our bindings
both BLAKE and \ac{BMW} outperform double \ac{SHA}-256 in every measurement,
with Luffa, Shabal and Simd outperforming double \ac{SHA}-256 in most of them.

We also measured single \ac{SHA}-256 for these files, which we believe should
show a difference. The lengths of inputs processed by the hash function are
dominated by sized of 106B for the 1B files, and also by 63B for the 1KB files,
as shown in Tables \ref{tbl:tahoe:datauploaded} and
\ref{tbl:tahoe:datadownloaded}. Even by the best-case performance scenario from
Table \ref{tbl:supercop:sha256d} we would expect to see a performance penalty.
However, this performance penalty is so small that it is typically hidden due to
system ``noise''.

\subsection{Large files}

For large files, 1MB, 100MB and 1GB, we can see that the \ac{SHA}-3
implementations are outperforming double \ac{SHA}-256. Also for the large files
we can see that there is a performance penalty introduced by our bindings (Table
\ref{tbl:cython:overhead}), however the relative overhead are small
compared to the time spent in the underlying libraries. The reason for
this is that we are hashing larger sizes of data, especially for the 100MB and
1GB files (Tables \ref{tbl:tahoe:datauploaded} and
\ref{tbl:tahoe:datauploaded}). For these larger sizes the overhead caused by
using double \ac{SHA}-256 opposed to single is also relatively small and does
not cause a noteworthy performance penalty.
